---
title: 'Peer-graded Assignment'
author: "swathy"
date: "5/7/2020"
output: html_document
---

## Loading the libraries

```{r libraries}
library(caret)
library(randomForest)
library(rpart)
```

## Loading the data

### Cross Validation(https://en.wikipedia.org/wiki/Cross-validation_(statistics))
I am going to partition the data(pml-training.csv) into training and testing by a factor of p<-0.65. I will then use the final prediction model of the testing dataset(pml-testing.csv)

```{r loading_data}
set.seed(12321)
data_training = read.csv(url('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'),na.strings=c("NA", ""))
data_testing = read.csv(url('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'),na.strings=c("NA", ""))
data_partition <- createDataPartition(data_training$classe, p=0.65, list=FALSE)
intrain <- data_training[data_partition,]
intest <- data_training[-data_partition,]
```

## Cleaning the data

```{r removing_NA}
intrain<-intrain[,colSums(is.na(intrain)) == 0]
intest <-intest[,colSums(is.na(intest)) == 0]
```

at this point we might want to remove all variables corresponding to first 7 variables, since they are not so important to the prediction

```{r removing_variables}
intrain<-intrain[,8:dim(intrain)[2]]
intest <-intest[,8:dim(intest)[2]]
```

## Predict using random forest

```{r predict_random_forest}
model_rf <- randomForest(classe ~. , data=intrain)
prediction_rf <- predict(model_rf, intest)
confusionMatrix(prediction_rf, intest$classe)
```

## Predict using Decision Trees

```{r predict_decision_trees}
model_dt <- rpart(classe ~. , data=intrain, method="class")
prediction_dt <- predict(model_dt, intest, type = "class")
confusionMatrix(prediction_dt, intest$classe)
```

## Conclusion of prediction

We see the accuracy of using Random Forest is better, as compared to decision trees. Hence we are going to use random forest to predict on the testing data.

```{r model}
model_rf
```


# Submission

```{r predict_final_testing_data}
prediction_dt <- predict(model_rf, data_testing)
prediction_dt
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



